---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r,echo=FALSE,results='hide',warning=FALSE}
Sys.setlocale("LC_TIME", "English")
library(ggplot2)
```

## Loading and preprocessing the data

Load the data from CSV file
```{r}
data <- read.csv("activity.csv", stringsAsFactors=FALSE)
str(data)
```

Parse strings as date
```{r}
data$date <- as.Date(data$date)
```

Show summary of the data
```{r}
summary(data)
```

## What is mean total number of steps taken per day?

Sum the number of steps taken each day
```{r}
stepsPerDay <- aggregate(data$steps, by=list(day = data$date), FUN=sum)
```

Draw a histogram of the values
```{r}
hist(stepsPerDay$x, breaks=20, col='red', main="Histogram of steps per day", xlab="Steps per day")
```

Calculate the mean and median values
```{r}
mean <- mean(stepsPerDay$x, na.rm=TRUE)
median <- median(stepsPerDay$x, na.rm=TRUE)
```
The mean value is `r mean`
The median value is `r median`

## What is the average daily activity pattern?

Average the number of steps taken during each interval for all days
```{r}
dailyAverage = aggregate(data$steps, by=list(interval = data$interval), FUN=mean, na.rm=TRUE)
```

Plot the daily activity pattern
```{r}
plot(dailyAverage$interval, dailyAverage$x, type='l', xlab="Interval", ylab="Average steps", main="Daily activity pattern")
```

Find max average value across day
```{r}
max <- max(dailyAverage$x)
maxInterval <- dailyAverage$interval[which.max(dailyAverage$x)]
```
The max average value is `r max`, it occurs for interval `r maxInterval`

## Imputing missing values

Calculate the number of missing values
```{r}
naIntervals <- sum(is.na(data$steps))
naDays <- sum(is.na(stepsPerDay$x))
```
There are total `r naIntervals` intervals with missing values and `r naDays` whole days with missing values.

Filling the missing values with mean value calculated from other days:
```{r}
dataFilled <- data
for(i in 1:length(data$steps)){
  if(is.na(data$steps[i])){
    dataFilled$steps[i] = dailyAverage$x[i]
  }
}
```

Sum the number of steps taken each day
```{r}
stepsPerDayFilled <- aggregate(dataFilled$steps, by=list(day = data$date), FUN=sum)
```

Draw a histogram of the values
```{r}
hist(stepsPerDayFilled$x, breaks=20, col='red', main="Histogram of steps per day with imputed missing variables", xlab="Steps per day")
```

Calculate the mean and median values
```{r}
meanFilled <- mean(stepsPerDayFilled$x, na.rm=TRUE)
medianFilled <- median(stepsPerDayFilled$x, na.rm=TRUE)
```
The mean value is `r meanFilled`
The median value is `r medianFilled`

The results differ, but not very much.

## Are there differences in activity patterns between weekdays and weekends?

Add weekday/weekend factor to original data
```{r}
data$day <- as.factor(ifelse(grepl("Saturday|Sunday",weekdays(data$date)), "weekend", "weekday"))
```

Aggregate the data by interval and day type
```{r}
weekDailyAverage <- aggregate(data$steps, by=list(interval = data$interval, day = data$day), FUN=mean, na.rm=TRUE)

```

Plot the daily activity patterns for weekends and weekdays
```{r}
qplot(interval, x,  data  =  weekDailyAverage,  facets = day~., xlab="Interval", ylab="Number of steps") + geom_line()
```








